Sketch of DP - Bayesian Inference

### 2018-07-17 main-sketch.tex
section 3 pass
second pass on section 3

\begin{abstract}
Bayesian inference is a statistical method which allows to derive a
\emph{posterior} distribution over a parameter, starting from a
\emph{prior} distribution and observed data.  In a setting where the
data must be protected, different approaches have been taken to make
this process differentially private. \citet{dimitrakakis2014robust},
and \citet{wang2015privacy}, for instance, proved that, under specific
conditions, sampling from the posterior distribution is already
differentially private. Other works,
e.g. \cite{zhang2016differential,foulds2016theory}, designed
differentially private mechanisms that output a representation of the
full posterior distribution. Also, accuracy of the mechanisms was
measured using metrics on the space of the numeric parameters of the
posterior distribution, e.g. $\ell_1$-norm .

@@

\section{Bayesian Inference Background}
\label{sec_background}
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and given an observation $\dataobs$,
the posterior distribution on $\theta$ given $\dataobs$ is computed as:
\[
  \Pr(\theta | \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression  $\Pr(\dataobs | \theta)$ denotes the \emph{likelihood} of $\theta$ when
$\dataobs$ is observed. Since we consider $\dataobs$ to be fixed the likelihood is a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$. That is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distributions. In this case we say that the prior
is conjugate of the likelihood function. Using conjugate priors simplify calculcations and allows
for inference to be performed in a recursive fashion over the data.
In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
Specifically, we will consider the situation where $\theta$ represents the parameter --informally called \emph{bias}--
of a Bernoulli distributed random variable, and its immediate generalization where the parameter $\vtheta$ represents
the the vector of parameters of a categorical distributed random variable.
In the former case the prior distribution over $\theta\in [0,1]$ is going to be a beta distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ 0/1 values, that is $\dataobs=\ll x_1,\dots x_n \gg, x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}x_i$.
From this it can easily be derived that the posterior distribution is:
\[
  \Pr(\theta|\dataobs)=\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)
\]
In the latter case the prior distribution over $\vtheta\in [0,1]^{k}$ is given by a Dirichelet distribution, $\dirichlet(\valpha)$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbf{R^{+}})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generealised beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
where $\Delta \alpha_i=\displaystyle\sum_{i=1}^{n}\iverson{x_i=a_i}$, where in $\iverson{\cdot}$ we use
Iverson bracket notation. The posterior distribution over $\vtheta$ turns out to be
\[
  \Pr(\vtheta|\dataobs)=\dirichlet(\valpha+\Delta \valpha).
\]
%where $+$ denotes the

\section{The Problem and Baseline approach}
We are interested in designing a mechanism for privately releasing
\emph{fully} the posterior distributions derived in section \ref{sec_background},
and not a sample from them.  It's worth noticing that the posterior
distributions are fully characterized by their parameters, and the
family (beta, Dirichlet) they belong to.  Hence, in case of the
Beta-Binomial model we are interested in releasing a private version
of the pair of parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)$, and
in the case of the Dirichlet-Multinomial model we are interested in a
private version of $\valpha'=(\valpha + \Delta \valpha)$. \todo{ \cite{?} and \cite{?}} have
already attacked this problem and their solution consisted in adding
independent Laplacian noise to the parameters of the posteriors. That
is, in the case of the Beta-Binomial system the value released would
be: $(\alpha + \Delta \tilde\alpha,\beta + n - \Delta \tilde\alpha)$ where $\Delta \tilde\alpha\sim Lap(\Delta \alpha, \frac{2}{\epsilon})$.
%with truncating method come from algorithm designed by \citet{zhang2016differential}
Where $Lap(\mu,\nu)$
denotes a Laplacian random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent\footnote{Given $\dataobs, \dataobs'\in\{0,1\}^{n}$ we say that $\dataobs$ and $\dataobs'$ are adjacent iff
$\displaystyle \sum_{i}^{n}\iverson{x_i = x'_i }\leq 1$.} data observations
$\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system.

@@

\begin{equation*}
\underset{z \thicksim \hexpmech}{\Pr}[z=r] = \frac {exp(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdotS_\beta(\dataobs)})}
{\displaystyle\sum_{r\in\betaset} exp(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdotS_{\beta}(\dataobs)})}
\end{equation*}
The previous mechanism is very similar to the exponential mechanism as defined in \cite{expmech},  with $\betaset$ as range and $\hellinger(\cdot,\cdot)$ as scoring function.
The difference is that in this mechanism we don't calibrate the noise w.r.t to the global sensitivity of the scoring function but w.r.t to the smooth sensitivity
-- \citet{nissim2007smooth}-- of $\hellinger(\bysinfer(\dataobs), \cdot)$, which is :
\[
   S_{\beta}(\dataobs)=\max_{\dataobs' \neq \dataobs, \dataobs' \in \{0,1\}^{n}}\bigg \{\Delta_{l}\cdot \hellinger(\bysinfer(\dataobs'),-)\cdot e^{-\gamma\cdot d(\dataobs, \dataobs')}\bigg\}
 \]

where $d$ is the Hamming distance between two datasets, $\Delta_{l}=????$, and $\gamma =
\gamma(\epsilon, \delta)$ is a function of $\epsilon$ and $\delta$, to be determined later.
\subsection{Analysis of the mechanism}
In what follows, we will use a correspondence between the probability
 $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$ of every
 $r\in\betaset$ and the probability 
 $\underset{z \thicksim \hexpmech(x)}{Pr}[\hlg(\bysinfer(x),z) =
 \hlg(\bysinfer(x),r)]$ for the utility score for $r$. In particular, for every
 $r\in\betaset$ we have:
$$
\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]=
\frac{1}{2}\Big (\underset{z \thicksim \hexpmech(x)}{Pr}[\hlg(\bysinfer(x),z) =
 \hlg(\bysinfer(x),r)]\Big )
$$
To see this, it is enough to notice that: $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$ is proportional too $\hlg(\bysinfer(x),r)$, i.e., $u(x,z)$. We can derive, if $u(r,x) = u(r',x)$ then $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r] = \underset{z \thicksim \hexpmech(x)}{Pr}[z = r']$. We assume the number of candidates $z \in \mathcal{R}$ that satisfy $u(z,x) = u(r,x)$ is $|r|$, we have  $\underset{z \thicksim \hexpmech(x)}{Pr}[u(z,x) = u(r,x)] = |r| \underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$. Because Hellinger distance  $\hlg(\bysinfer(x),z)$ is axial symmetry, where the $\bysinfer(x)$ is the symmetry axis. It can be infer that $|z| = 2$ for any candidates, apart from the true output, i.e., $\underset{z \thicksim \hexpmech(x)}{Pr}[u(z,x) = u(r,x)] = 2 \underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$. This parameter can be eliminate in both sides in proof.

>>>>

\begin{abstract}
Bayesian inference is a statistical method which allows to derive a
\emph{posterior} distribution over a parameter, starting from a
\emph{prior} distribution and observed data.  In a setting where the
data must be protected, different approaches have been taken to make
this process differentially private. \citet{dimitrakakis2014robust},
and \citet{wang2015privacy}, for instance, proved that, under specific
conditions, sampling from the posterior distribution is already
differentially private. Other works,
e.g. \cite{zhang2016differential}, \cite{foulds2016theory}, designed
differentially private mechanisms that output a representation of the
full posterior distribution. Also, accuracy of the mechanisms was
measured using metrics on the space of the numeric parameters of the
posterior distribution, e.g. $\ell_1$-norm .

@@

\section{Bayesian Inference Background}
\label{sec_background}
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and given an observation $\dataobs$,
the posterior distribution on $\theta$ given $\dataobs$ is computed as:
\[
  \Pr(\theta | \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression  $\Pr(\dataobs | \theta)$ denotes the \emph{likelihood} of $\theta$ when
$\dataobs$ is observed. Since we consider $\dataobs$ to be fixed the likelihood is a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$. That is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distributions. In this case we say that the prior
is conjugate of the likelihood function. Using conjugate priors simplify calculcations and allows
for inference to be performed in a recursive fashion over the data.
In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
Specifically, we will consider the situation where $\theta$ represents the parameter --informally called \emph{bias}--
of a Bernoulli distributed random variable, and its immediate generalization where the parameter $\vtheta$ represents
the the vector of parameters of a categorical distributed random variable.
In the former case the prior distribution over $\theta\in [0,1]$ is going to be a beta distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ 0/1 values, that is $\dataobs=\ll x_1,\dots x_n \gg, x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}x_i$.
From this it can easily be derived that the posterior distribution is:
\[
  \Pr(\theta|\dataobs)=\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)
\]
In the latter case the prior distribution over $\vtheta\in [0,1]^{k}$ is given by a Dirichelet distribution, $\dirichlet(\valpha)$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbf{R^{+}})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generealised beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
where $\Delta \alpha_i=\displaystyle\sum_{i=1}^{n}\iverson{x_i=a_i}$, where in $\iverson{\cdot}$ we use
Iverson bracket notation. The posterior distribution over $\vtheta$ turns out to be
\[
  \Pr(\vtheta|\dataobs)=\dirichlet(\valpha+\Delta \valpha). 
\]
where $+$ denotes the component wise  sum of vectors of reals.

\section{The Problem and Baseline approach}
We are interested in designing a mechanism for privately releasing
\emph{fully} the posterior distributions derived in section \ref{sec_background},
and not a sample from them.  It's worth noticing that the posterior
distributions are fully characterized by their parameters, and the
family (beta, Dirichlet) they belong to.  Hence, in case of the
Beta-Binomial model we are interested in releasing a private version
of the pair of parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)$, and
in the case of the Dirichlet-Multinomial model we are interested in a
private version of $\valpha'=(\valpha + \Delta \valpha)$. \citet{zhang2016differential} and \citet{xiao2012bayesian} have
already attacked this problem and their solution consisted in adding
independent Laplacian noise to the parameters of the posteriors. That
is, in the case of the Beta-Binomial system the value released would
be: $(\tilde\alpha,\tilde\beta)=(\alpha + \Delta \tilde\alpha,\beta + n - \Delta \tilde\alpha)$ where $\Delta \tilde\alpha\sim Lap(\Delta \alpha, \frac{2}{\epsilon})$, and
%with truncating method come from algorithm designed by \citet{zhang2016differential}
where $Lap(\mu,\nu)$
denotes a Laplacian random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent\footnote{Given $\dataobs, \dataobs'\in\{0,1\}^{n}$ we say that $\dataobs$ and $\dataobs'$ are adjacent and we write, $\adj{\dataobs}{\dataobs'}$, iff
$\displaystyle \sum_{i}^{n}\iverson{x_i = x'_i }\leq 1$ } data observations
$\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system.

@@

\begin{equation*}
\underset{z \thicksim \hexpmech}{\Pr}[z=r] = \frac {exp(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdot S_\beta(\dataobs)})}
{\displaystyle\sum_{r\in\betaset} exp(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdot S_{\beta}(\dataobs)})}
\end{equation*}
The previous mechanism is very similar to the exponential mechanism as defined in \cite{expmech},  with $\betaset$ as range and $\hellinger(\cdot,\cdot)$ as scoring function.
The difference is that in this mechanism we don't calibrate the noise w.r.t to the global sensitivity of the scoring function but w.r.t to the smooth sensitivity
-- \citet{nissim2007smooth}-- of $\hellinger(\bysinfer(\dataobs), \cdot)$, which is :
\[
   S_{\beta}(\dataobs)=\max_{\dataobs' \neq \dataobs, \dataobs' \in \{0,1\}^{n}}\bigg \{\Delta_{l}\bigg (\hellinger(\bysinfer(\dataobs'),-)\bigg )\cdot e^{-\gamma\cdot d(\dataobs, \dataobs')}\bigg\}
 \]

 where $d$ is the Hamming distance between two datasets,  $\gamma = \gamma(\epsilon, \delta)$ is a function
 of $\epsilon$ and $\delta$ to be determined later, and where $\Delta_{l}\bigg (\hellinger(\bysinfer(\dataobs'),\cdot)\bigg )$
 denotes the local sensitivity at $\dataobs'$ (w.r.t $\ell_1$ norm)
 of the function
 \[\hellinger \circ \langle (\bysinfer(\dataobs', \cdot), \bysinfer(\cdot, \cdot)\rangle\]
 that is:
 \begin{equation*}
 \max_{\dataobs'' \in \datauni^n:\adj{\dataobs'}{\dataobs''}, r\in \betaset}\lvert \hellinger(\bysinfer(\dataobs'), r) - \hellinger(\bysinfer(\dataobs''), r)\rvert
\end{equation*}
Notice that we maximize also over $r$\todo{need to explain why}.
\subsection{Analysis of the mechanism}
In what follows, we will use a correspondence between the probability
 $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$ of every
 $r\in\betaset$ and the probability 
 $\underset{z \thicksim \hexpmech(x)}{Pr}[\hlg(\bysinfer(x),z) =
 \hlg(\bysinfer(x),r)]$ for the utility score for $r$. In particular, for every
 $r\in\betaset$ we have:
$$
\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]=
\frac{1}{2}\Big (\underset{z \thicksim \hexpmech(x)}{Pr}[\hlg(\bysinfer(x),z) =
 \hlg(\bysinfer(x),r)]\Big )
$$
To see this, it is enough to notice that: $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$ is proportional too $\hlg(\bysinfer(x),r)$, i.e., $u(x,z)$. We can derive, if $u(r,x) = u(r',x)$ then $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r] = \underset{z \thicksim \hexpmech(x)}{Pr}[z = r']$. We assume the number of candidates $z \in \mathcal{R}$ that satisfy $u(z,x) = u(r,x)$ is $|r|$, we have  $\underset{z \thicksim \hexpmech(x)}{Pr}[u(z,x) = u(r,x)] = |r| \underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$. Because Hellinger distance  $\hlg(\bysinfer(x),z)$ is axial symmetry, where the $\bysinfer(x)$ is the symmetry axis. It can be infer that $|z| = 2$ for any candidates, apart from the true output, i.e., $\underset{z \thicksim \hexpmech(x)}{Pr}[u(z,x) = u(r,x)] = 2 \underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$. This parameter can be eliminate in both sides in proof.
